---
title      : "Analysis -  Descriptive Analysis"
editor     : visual
format     : docx
prefer-html: true
echo       : false
date       : "March 8, 2023"
---

### Purpose

Calculate descriptive statistics for general and military health in both tabular and graphical formats.

### Setup {.unnumbered .unlisted}

```{r}
#| warning: false
#| message: false
#| echo:    false

# ----------------------------------- #
# Clear working environment  
# ----------------------------------- #
rm(list = ls())
# ----------------------------------- #

# ----------------------------------- #
# Load required packages
# ----------------------------------- #
library(tidyverse)
library(sf)
library(kableExtra)
# library(countrycode)
# library(lubridate)
# ----------------------------------- #

# ----------------------------------- #
# Load data
# ----------------------------------- #
d <- read_rds("../../Data/analysis-data/0-analysis_data.rds")
# ----------------------------------- #
```

### Top-10 military health recipients

Calculate top recipients based on three metrics:

1.  Total number of projects
2.  Total amount spent
3.  Total amount spent per project (average)

Data coverage includes all projects between `r min(d$year)` and `r max(d$year)`.

```{r}
#| echo:    false

mil_top <- d %>%
  st_drop_geometry %>% 
  rename(mil_health_us_amount = usaid_amount_mil, 
         mil_health_cn_amount = aiddata_amount_mil) %>% 
  select(cname, 
         mil_health_us_count,  mil_health_cn_count,
         mil_health_us_amount, mil_health_cn_amount) %>% 
  group_by(cname) %>% 
  summarize(across(.cols = ends_with("count"), 
                   .fns  = sum),
            across(.cols = ends_with("amount"),
                   .fns  = ~(sum(.x, na.rm = TRUE) / 1e6))) %>% 
  ungroup %>% 
  mutate(
    mil_health_us_perproject = mil_health_us_amount / mil_health_us_count,
    mil_health_cn_perproject = mil_health_cn_amount / mil_health_cn_count
    ) %>% 
  mutate(
    mil_health_us_perproject = case_when(
      (mil_health_us_count == 0 & mil_health_us_amount == 0) ~ 0, TRUE ~ mil_health_us_perproject
    ),
    mil_health_cn_perproject = case_when(
      (mil_health_cn_count == 0 & mil_health_cn_amount == 0) ~ 0, TRUE ~ mil_health_cn_perproject
    )
  )
```

Note - all summary calculations include top-11 for US and top-10 for China to facilitate comparison with US including or excluding Iraq.

{{< pagebreak >}}

#### Top military - total projects

```{r}
#| echo:    false

mil_top_us <- mil_top %>% 
  select(cname, mil_health_us_count) %>%
  arrange(desc(mil_health_us_count)) %>% 
  slice(1:11) %>% 
  mutate(us = sprintf("%s - %s", cname, mil_health_us_count)) %>% 
  pull(us)

mil_top_cn <- mil_top %>% 
  select(cname, mil_health_cn_count) %>%
  arrange(desc(mil_health_cn_count)) %>% 
  slice(1:10) %>% 
  mutate(cn = sprintf("%s - %s", cname, mil_health_cn_count)) %>% 
  pull(cn)

mil_top_ct <- tibble(
  "United States" = mil_top_us, 
  "China"         = c(mil_top_cn, NA)
)
```

```{r}
#| echo:    false

kbl(mil_top_ct, caption = "Top military health recipients: total projects") %>% 
  # row_spec(0:11, bold = F, color = "black", background = "white") %>% 
    footnote(.,
           general = paste0(
             "Data sources: US - USAID, CN - AidData."
           ))
```

{{< pagebreak >}}

#### Top military - total spending

```{r}
#| echo:    false

mil_top_us <- mil_top %>% 
  select(cname, mil_health_us_amount) %>%
  arrange(desc(mil_health_us_amount)) %>% 
  slice(1:11) %>% 
  mutate(mil_health_us_amount = as.character(round(mil_health_us_amount, 2))) %>% 
  mutate(us = sprintf("%s - %s", cname, mil_health_us_amount)) %>% 
  pull(us)

mil_top_cn <- mil_top %>% 
  select(cname, mil_health_cn_amount) %>%
  arrange(desc(mil_health_cn_amount)) %>% 
  slice(1:10) %>% 
  mutate(mil_health_cn_amount = as.character(round(mil_health_cn_amount, 2))) %>% 
  mutate(cn = sprintf("%s - %s", cname, mil_health_cn_amount)) %>% 
  pull(cn)

mil_top_sp <- tibble(
  "United States" = mil_top_us, 
  "China"         = c(mil_top_cn, NA)
)
```

```{r}
#| echo:    false

kbl(mil_top_sp, 
    caption = "Top military health recipients: total spending") %>% 
  # row_spec(0:11, bold = F, color = "black", background = "white") %>% 
  footnote(.,
           general = paste0(
             "Amounts reported in 2017 constant USD, millions.\n",
             "Data sources: US - USAID, CN - AidData."
           ))
```

{{< pagebreak >}}

#### Top military - average spending per project

```{r}
#| echo:    false

mil_top_us <- mil_top %>% 
  select(cname, mil_health_us_perproject) %>%
  arrange(desc(mil_health_us_perproject)) %>% 
  slice(1:11) %>% 
  mutate(mil_health_us_perproject = as.character(round(mil_health_us_perproject, 2))) %>% 
  mutate(us = sprintf("%s - %s", cname, mil_health_us_perproject)) %>% 
  pull(us)

mil_top_cn <- mil_top %>% 
  select(cname, mil_health_cn_perproject) %>%
  arrange(desc(mil_health_cn_perproject)) %>% 
  slice(1:10) %>% 
  mutate(mil_health_cn_perproject = as.character(round(mil_health_cn_perproject, 2))) %>% 
  mutate(cn = sprintf("%s - %s", cname, mil_health_cn_perproject)) %>% 
  pull(cn)

mil_top_pp <- tibble(
  "United States" = mil_top_us, 
  "China"         = c(mil_top_cn, NA)
)
```

```{r}
#| echo:    false

kbl(mil_top_pp, caption = "Top military health recipients: average spending per project") %>% 
  # row_spec(0:11, bold = F, color = "black", background = "white") %>% 
  footnote(.,
           general = paste0(
             "Amounts reported in 2017 constant USD, millions per project.\n",
             "Data sources: US - USAID, CN - AidData."
           ))
```

{{< pagebreak >}}

### Summary statistics

#### Health outcomes

```{r}
#| echo:    false

y <- d %>% 
  
  st_drop_geometry %>% 
  select(dah_us, dah_us_per, 
         dah_cn, dah_cn_per,
         
         mil_health_us_count, mil_health_us, 
         mil_health_cn_count, mil_health_cn,
         
         cdist_us, un_gdppc, un_lifeexpec_yrs, un_infdeath_p1000, 
         ghdx_cmd_prevalence_pc, ghdx_ncd_prevalence_pc, 
         emdat_disasters_total, emdat_deaths_total, 
         ged_major, 
         atop_us_nonagg, atop_cn_nonagg, 
         vdem
         ) %>% 
  
  mutate(across(.cols = c(dah_us, dah_cn), 
                .fns  = ~(.x / 1e6))) %>% 
  
  summarize(across(.cols = everything(), 
                   .fns  = list("Mean"   = ~mean(.x,   na.rm = TRUE),
                                "Median" = ~median(.x, na.rm = TRUE),
                                "SD"     = ~sd(.x,     na.rm = TRUE),
                                "Min"    = ~min(.x,    na.rm = TRUE),
                                "Max"    = ~max(.x,    na.rm = TRUE),
                                "N"      = ~n()),
                   .names= "{.col}.{.fn}"))

y <- y %>% 
  pivot_longer(data = ., 
               cols    = everything(),
               names_to = c("var", "stat"), 
               names_sep = "[.]", 
               values_to = "val")

y <- y %>% 
  pivot_wider(., 
              id_cols = var, 
              names_from = stat, 
              values_from = val) %>% 
  mutate(vtype = case_when(
    str_starts(var, "dah|mil") ~ "outcome", 
    TRUE ~ "control"
  ))

y <- y %>% 
  mutate(Source = case_when(
    str_detect(var, "_cn") ~ "China",
    str_detect(var, "_us") ~ "United States"
    )) %>% 
  mutate(Measure = case_when(
    var %in% c("dah_us", "dah_cn")         ~ "Health spending (USD)", 
    var %in% c("dah_us_per", "dah_cn_per") ~ "Health allocation (% budget)",
    var %in% c("mil_health_us", "mil_health_cn") ~ "Military health (binary)",
    var %in% c("mil_health_us_count", "mil_health_cn_count") ~ "Military health (projects)",
    
    var == "cdist_us" ~ "Distance from US (km, log)",
    var == "un_gdppc" ~ "GDP per capita (log)",
    var == "un_lifeexpec_yrs" ~ "Life expectancy (years)",
    var == "un_infdeath_p1000" ~ "Infant death rate (per 1000 live births, log)",
    var == "ghdx_cmd_prevalence_pc" ~ "Communicable disease per capita (log)",
    var == "ghdx_ncd_prevalence_pc" ~ "Non-communicable disease per capita (log)",
    var == "emdat_disasters_total" ~ "Humanitarian crisis count",
    var == "emdat_deaths_total" ~ "Humanitarian crisis deaths (log)",
    var == "ged_major" ~ "Civil war",
    var == "atop_us_nonagg" ~ "Non-agression pact: US",
    var == "atop_cn_nonagg" ~ "Non-agression pact: CN",
    var == "vdem" ~ "Democracy"
    )) %>% 
  relocate(Source,  .before = 1) %>% 
  relocate(Measure, .after = Source) 
```

##### Health - General

```{r}
#| echo:    false

y %>% 
  filter(!str_detect(Measure, "Military"), vtype == "outcome") %>% 
  select(-vtype, -var) %>% 
  
  kbl(., caption = "Health - General: Summary statistics", digits = 3) %>% 
    collapse_rows(kable_input = ., 
                  columns     = 1,
                  valign      = "middle") %>% 
    footnote(kable_input = .,
           general = paste0(
             "USD measures reflect 2017 constant millions USD.\n",
             "Data source: IHME DAH project."
           )) 
```

##### Health - Military

```{r}
#| eval: false

table("CN" = d$mil_health_cn, "US" = d$mil_health_us) %>% 
  apply(., 1, function(x){x / sum(x)}) %>%
  t %>% round(., 2)
```

```{r}
#| echo:    false

y %>% filter(str_detect(Measure, "Military"), vtype == "outcome") %>% 
  select(-vtype, -var) %>% 

  kbl(., caption = "Health - Military: Summary statistics", digits = 3) %>% 
    collapse_rows(kable_input = ., 
                  columns     = 1,
                  valign      = "middle") %>% 
    footnote(kable_input = .,
           general = paste0(
             "Data sources: US - USAID, CN - AidData."
           ),
           threeparttable = TRUE) 
```

{{< pagebreak >}}

### Health outcomes (no Iraq)

Note - main model results are robust to data including or excluding Iraq. Many of the 298 Iraq projects in top-reported year (XXXX) are Commander's Emergency Response Program allocations.

```{r}
#| echo:    false

y2 <- d %>% 
  
  filter(cname != "Iraq") %>% 
  
  st_drop_geometry %>% 
  select(dah_us, dah_us_per, 
         dah_cn, dah_cn_per,
         
         mil_health_us_count, mil_health_us, 
         mil_health_cn_count, mil_health_cn,
         
         cdist_us, un_gdppc, un_lifeexpec_yrs, un_infdeath_p1000, 
         ghdx_cmd_prevalence_pc, ghdx_ncd_prevalence_pc, 
         emdat_disasters_total, emdat_deaths_total, 
         ged_major, 
         atop_us_nonagg, atop_cn_nonagg, 
         vdem
         ) %>% 
  
  mutate(across(.cols = c(dah_us, dah_cn), 
                .fns  = ~(.x / 1e6))) %>% 
  
  summarize(across(.cols = everything(), 
                   .fns  = list("Mean"   = ~mean(.x,   na.rm = TRUE),
                                "Median" = ~median(.x, na.rm = TRUE),
                                "SD"     = ~sd(.x,     na.rm = TRUE),
                                "Min"    = ~min(.x,    na.rm = TRUE),
                                "Max"    = ~max(.x,    na.rm = TRUE),
                                "N"      = ~n()),
                   .names= "{.col}.{.fn}"))

y2 <- y2 %>% 
  pivot_longer(data = ., 
               cols    = everything(),
               names_to = c("var", "stat"), 
               names_sep = "[.]", 
               values_to = "val")

y2 <- y2 %>% 
  pivot_wider(., 
              id_cols = var, 
              names_from = stat, 
              values_from = val) %>% 
  mutate(vtype = case_when(
    str_starts(var, "dah|mil") ~ "outcome", 
    TRUE ~ "control"
  ))

y2 <- y2 %>% 
  mutate(Source = case_when(
    str_detect(var, "_cn") ~ "China",
    str_detect(var, "_us") ~ "United States"
    )) %>% 
  mutate(Measure = case_when(
    var %in% c("dah_us", "dah_cn")         ~ "Health spending (USD)", 
    var %in% c("dah_us_per", "dah_cn_per") ~ "Health allocation (% budget)",
    var %in% c("mil_health_us", "mil_health_cn") ~ "Military health (binary)",
    var %in% c("mil_health_us_count", "mil_health_cn_count") ~ "Military health (projects)",
    
    var == "cdist_us" ~ "Distance from US (km, log)",
    var == "un_gdppc" ~ "GDP per capita (log)",
    var == "un_lifeexpec_yrs" ~ "Life expectancy (years)",
    var == "un_infdeath_p1000" ~ "Infant death rate (per 1000 live births, log)",
    var == "ghdx_cmd_prevalence_pc" ~ "Communicable disease per capita (log)",
    var == "ghdx_ncd_prevalence_pc" ~ "Non-communicable disease per capita (log)",
    var == "emdat_disasters_total" ~ "Humanitarian crisis count",
    var == "emdat_deaths_total" ~ "Humanitarian crisis deaths (log)",
    var == "ged_major" ~ "Civil war",
    var == "atop_us_nonagg" ~ "Non-agression pact: US",
    var == "atop_cn_nonagg" ~ "Non-agression pact: CN",
    var == "vdem" ~ "Democracy"
    )) %>% 
  relocate(Source,  .before = 1) %>% 
  relocate(Measure, .after = Source) 
```

##### Health - General (no Iraq)

```{r}
#| echo:    false

y2 %>% filter(!str_detect(Measure, "Military"), vtype == "outcome") %>% 
  select(-vtype, -var) %>% 

  kbl(., caption = "Health - General: Summary statistics", digits = 3) %>% 
    collapse_rows(kable_input = ., 
                  columns     = 1,
                  valign      = "middle") %>% 
    footnote(kable_input = .,
           general = paste0(
             "USD measures reflect 2017 constant millions USD.\n",
             "Data source: IHME DAH project."
           )) 
```

##### Health - Military (no Iraq)

```{r}
#| echo:    false

y2 %>% filter(str_detect(Measure, "Military"), vtype == "outcome") %>% 
  select(-vtype, -var) %>% 

  kbl(., caption = "Health - Military: Summary statistics", digits = 3) %>% 
    collapse_rows(kable_input = ., 
                  columns     = 1,
                  valign      = "middle") %>% 
    footnote(kable_input = .,
           general = paste0(
             "Data sources: US - USAID, CN - AidData."
           )) 
```

{{< pagebreak >}}

### Control variables

```{r}
y %>% filter(vtype == "control") %>% 
    select(-vtype, -var, - Source) %>% 

  kbl(., caption = "Control variables: Summary statistics", digits = 3) %>% 
    collapse_rows(kable_input = ., 
                  columns     = 1,
                  valign      = "middle") %>% 
    footnote(kable_input = .,
           general = "Data sources:",
           number  = c("Distance - locally computed in GIS",
                       "UN Stats: GDP, life expec., infant death", 
                       "GHDX: Communicable and non-communicable disease prevalence",
                       "EM-Dat: Humanitarian crises and deaths",
                       "UCDP GED: Civil war", 
                       "ATOP: non-aggression pacts",
                       "VDEM: democracy")) 
```
