---
title : "Analysis -  Military health"
editor: visual
date  : "March 8, 2023"
---

## Purpose

Estimate military health aid models.

## Setup

```{r}
#| warning: false
#| message: false
# ----------------------------------- #
# Clear working environment  
# ----------------------------------- #
rm(list = ls())
# ----------------------------------- #

# ----------------------------------- #
# Load required packages
# ----------------------------------- #
library(tidyverse)
library(sf)
library(texreg)
library(sandwich)
library(lmtest)
library(margins)
library(kableExtra)
# ----------------------------------- #

# ----------------------------------- #
# Load data
# ----------------------------------- #
d <- read_rds("../../Data/analysis-data/0-analysis_data.rds")
# ----------------------------------- #
```

## Military health

```{r}

# COMMENTED OUT IS LIKELY COORRECT WAY
# In case I messed up the polynomial.
d <- d %>%
  st_drop_geometry %>%
  arrange(ccode, year) %>%
  group_by(ccode) %>%
  mutate(change = mil_health_us != mil_health_us_lag) %>%
  mutate(poly1 = sequence(table(cumsum(change) + 1)),
         poly2 = poly1^2,
         poly3 = poly1^3)
```

```{r}
# d <- d %>%
#   arrange(as.numeric(ccode), as.numeric(year)) %>% 
#   group_by(ccode) %>% 
#   mutate(poly1 = 1:n(),
#          poly2 = poly1^2,
#          poly3 = poly1^3) %>% 
#   ungroup
```

### Formula

```{r}
# fs <- list(
#   "f1" = . ~ poly1 + poly2 + poly3 + region,
#   "f2" = . ~ 
#     # cdist_us +
#     un_gdppc +
#     un_gdppc2 +
#     un_lifeexpec_yrs +
#     # un_lifeexpec_yrs2 +
#     un_infdeath_p1000 +
#     ghdx_ncd_prevalence_pc + ghdx_cmd_prevalence_pc +
#     emdat_disasters_total + emdat_deaths_total +
#     ged_major +
#     atop_nonagg +
#     vdem + I(vdem^2) +
#     poly1 + poly2 + poly3 + 
#     region
# )

base    <-  . ~ poly1 + poly2 + poly3 + region
full    <-  update(base, ~ . + un_gdppc + un_gdppc2 + 
                     un_lifeexpec_yrs + un_lifeexpec_yrs2 + un_infdeath_p1000 + 
                     ghdx_ncd_prevalence_pc + ghdx_cmd_prevalence_pc + 
                     emdat_disasters_peryr + ged_major + atop_nonagg + 
                     trade_net + vdem + vdem2)

fs <- list(base = base, full = full)
rm(base, full)
```

### Fit

```{r}
us <- lapply(fs, function(x){
  glm(formula = update(x, mil_health_us ~ mil_health_cn + .),
      data    = d, 
      family  = binomial(link = "probit"))
})

cn <- lapply(fs, function(x){
  glm(formula = update(x, mil_health_cn ~ mil_health_us + .),
      data    = d, 
      family  = binomial(link = "probit"))
})

# Models list
mods_military <- list("us" = us, "cn" = cn) %>% list_flatten()

# Goodness of fit statistics
mods_military_gof <- list(
  "FE Region"      = rep("Yes", length(mods_military)),
  "Time polynomial"= rep("Yes", length(mods_military)),
  "Obs."           = lapply(mods_military, function(x) nrow(model.matrix(x))),
  "Log Likelihood" = lapply(mods_military, function(x) round(logLik(x) %>% as.numeric(), 2))
)

# rm(us, cn)
```

## DYNAMIC EFFECTS PER CARTER AND SIGNORINO

-   NOTE -- figure out which spline is right ! Read the paper - the two splines I have are producing distinct curves.
-   SEE `tmp-probit-dynamics.R`

### Robust SEs

```{r}
mods_military_robust <- lapply(mods_military, function(x){
  vcv <- vcovPC(x = x, cluster = ~ ccode + year, pairwise = F)
  # vcv <- vcovCL(x = x, cluster = ~ ccode + year)
  # vcv <- vcovHC(x = x, type = "HC0")
  # mrb <- coeftest(x = x, vcov. = vcv)
  return(vcv)
})
```

### Table

```{r}
nmod <- length(mods_military)
mm <- lapply(names(mods_military), function(x){return(
  coeftest(x = mods_military[[x]], 
           vcov. = mods_military_robust[[x]])
)})
screenreg(mm,
          digits = 3,
          # custom.coef.map = var_names,
          custom.header = list("US" = 1:(nmod/2), "CN" = ((nmod/2)+1):nmod),
          stars = c(0.001, 0.01, 0.05, 0.1),
          omit.coef = "year|region|cname|ccode",

          custom.gof.rows = mods_military_gof

          # custom.note = cnote
          )
```

------------------------------------------------------------------------

# \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## Quantities of interest

### Military probabilities

```{r}
p_cn_on_us <- margins(model     = mods_military$us_full, 
                      variables = "mil_health_cn", type = "response",
                      vcov      = mods_military_robust$us_base)
p_us_on_cn <- margins(model     = mods_military$cn_full, 
                      variables = "mil_health_us", type = "response",
                      vcov      = mods_military_robust$cn_full)
```

```{r}
p_cn_on_us <- summary(p_cn_on_us, level = 0.9) %>% as_tibble
p_us_on_cn <- summary(p_us_on_cn, level = 0.9) %>% as_tibble

mods_military_qoi <- bind_rows(p_cn_on_us, p_us_on_cn) %>% 
  dplyr::select(factor, AME, lower, upper) %>% 
  mutate(across(.cols = !factor, .fns = ~(.x * 100))) %>% 
  mutate(
    x    = c(1, 1.5),
    xlab = c("US", "CN"),
    clab = c("CN military presence", "US military health presence")
  )
```

```{r}
mil_probs <- ggplot(mods_military_qoi, aes(x = x)) +
  geom_errorbar(aes(ymin = lower, ymax = upper, color = clab),
                width = 0.1) +
  geom_point(aes(y = AME, color = clab)) +
  geom_hline(aes(yintercept = 0), color = "black", linewidth = 0.3) +
  theme(
    panel.background   = element_rect(fill = NA, color = "black"),
    panel.grid.major.y = element_line(color= "gray80", linetype = "dotted", linewidth = 0.25),
    axis.title.x       = element_blank(),
    legend.position    = "bottom"
  ) + 
  scale_x_continuous(breaks = mods_military_qoi$x, labels = mods_military_qoi$xlab) + 
  # lims(x = c(0.75, 1.75), y = c(-1, 15)) + 
  # coord_fixed(ratio = .1, xlim = c(0.75, 1.75), ylim = c(-1, 15), expand = T) +
  guides(color = guide_legend(title = "Effect of:", nrow = 2, title.position = "top")) + 
  labs(
    title = "Predicted probability of military health presense",
    subtitle = "in response to military health presenece of other power")

# ggsave(plot = mil_probs, 
#        filename = "../../Results/Figures/3-health_military-probabilities.png",
#        device = "png",
#        width    = 5.0,
#        height   = 5.0, 
#        units    = "in")

png(filename = "../../Results/Figures/3-health_military-probabilities.png",
    width    = 5, 
    height   = 5,
    units    = "in",
    res      = 360)
mil_probs
dev.off()

rm(mil_probs)
```

### Save

```{r}
save(mods_military, mods_military_gof, mods_military_qoi, 
     file = "../../Data/analysis-data/3-models_health-military.Rdata")

# rm(list = ls())
```

------------------------------------------------------------------------

# \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## Robustness checks

-   Additional models to explore robustness of results

Note - this subsetting block tests robustness of results by retaining only those countries in the data which have at least 1 year of either US or Chinese military health aid. Main results retain their significance at the 0.1 level.

### Subset - at least 1 year mil aid

```{r}
dsub <- d %>%
  st_drop_geometry() %>%
  group_by(cname) %>%
  summarize(tgt = ifelse(any(mil_health_cn == 1) | any(mil_health_us == 1), 1, 0)) %>%
  ungroup %>%
  filter(tgt == 1) %>%
  pull(cname)

d2 <- d %>%
  filter(cname %in% dsub)
```

### SAR Probits

```{r}
#| eval: false

library(spatialprobit)

load("Data/analysis-data/0-spatial_weights.rdata")
w <- Matrix(d_spatial$w_knn, sparse = TRUE)

?sarprobit
us <- sarprobit(formula = update(fs$full, mil_health_us ~ mil_health_us_lag + 
                                   mil_health_cn + . - region),
                W       = w,
                data    = d,
                ndraw   = 2e3,
                burn.in = 1e3)

summary(us)

cn <- sarprobit(formula = update(fs$full, mil_health_cn ~ mil_health_cn_lag + 
                                   mil_health_us + . - region),
                W       = w,
                data    = d,
                ndraw   = 2e3,
                burn.in = 1e3)

summary(cn)
```

Concerning military model fit:

-   With contemporaneous rival -- US and China significant
-   With contemporaneous rival + lag rival - Only US significant
-   With contemporaneous rival + lag rival + own lag - Only US significant
-   With contemporaneous rival + own lag - **Both significant**
    -   NOTE - not robust to inclusion of year FE - but probably overkill with time poly.

    -   Plus - based on error appears separation occurs with year FE.
