---
title : "Analysis - Effects, military"
editor: visual
date  : "March 28, 2023"
---

# NOTES

1.  implement dnorm() derivative method to calculate marginal effects rather than margins()
    1.  see tmp-probit-effs-20230328.R
2.  calculate inflection point for quadratic terms

## Purpose

Calculate marginal effects of military health models

## Setup

```{r}
#| warning: false
#| message: false
# ----------------------------------- #
# Clear working environment  
# ----------------------------------- #
rm(list = ls())
# ----------------------------------- #

# ----------------------------------- #
# Load required packages
# ----------------------------------- #
library(tidyverse)
library(sf)
library(margins)
library(kableExtra)
# ----------------------------------- #

# ----------------------------------- #
# Load data
# ----------------------------------- #
d <- read_rds("../../Data/analysis-data/0-analysis_data.rds")
load("../../Data/analysis-data/3-models_health-military.Rdata")
# ----------------------------------- #

# ----------------------------------- #
# Functions
# ----------------------------------- #
# source("../Functions/dynamics.R")
# rm(conditional_eff, half_life, lrss_star, lrss_time,
#    mr_star, mr_star_ev, mr_time)
source("../Functions/themes.R")
# ----------------------------------- #
```

### Create storage list

```{r}
military_effects <- list("tables"  = list(), 
                         "figures" = list())
```

## Marginal effects - main results

-   Inclusive of cubic polynomial terms - no proportional hazards over time.

```{r}
# dyn <- lapply(c("us_full", "cn_full"), function(x){
#   probit_response(m = mods_military[[x]], vcv = mods_military_robust[[x]], sims = 1000) %>% 
#     mutate(model := x)
# }) %>% bind_rows()
#
# me_us <- dyn %>% filter(vr == "fd", model == "us_full") %>% 
#   select(lb:ub) %>% apply(., 2, median)
# 
# me_cn <- dyn %>% filter(vr == "fd", model == "cn_full") %>% 
#   select(lb:ub) %>% apply(., 2, median)
# 
# dyn_eff <- bind_rows(me_us, me_cn) %>% 
#   mutate(model = c("us", "cn"), y = c(1, 2))
# 
# rm(me_us, me_cn, dyn)
#
# ceiling_dec <- function(x, level=1) round(x + 5*10^(-level-1), level)
# 
# ggplot(data = dyn_eff, aes(y = y, color = model)) + 
#   geom_errorbar(aes(xmin = lb, xmax = ub), width = 0.1) + 
#   geom_point(aes(x = md)) + 
#   scale_x_continuous(labels = function(x){ x * 100}, 
#                      limits = c(0, ceiling_dec(max(dyn_eff$ub))))
```

### Compute margins

```{r}
p_cn_on_us <- margins(model     = mods_military$us_full, 
                      variables = "mil_health_cn", type = "response",
                      vcov      = mods_military_robust$us_full)
p_us_on_cn <- margins(model     = mods_military$cn_full, 
                      variables = "mil_health_us", type = "response",
                      vcov      = mods_military_robust$cn_full)

p_cn_on_us <- summary(p_cn_on_us, level = 0.95) %>% as_tibble
p_us_on_cn <- summary(p_us_on_cn, level = 0.95) %>% as_tibble

eff_mil <- bind_rows(p_us_on_cn %>% select(AME, lower, upper) %>% mutate(model = "cn"),
                     p_cn_on_us %>% select(AME, lower, upper) %>% mutate(model = "us")) %>% 
  rename(md = AME, lb = lower, ub = upper) %>% 
  mutate(y       = c(1, 2), 
         y_label = c("CN response\nto US", "US response\nto CN")) %>% 
  select(y_label, lb, md, ub, model, y)

rm(p_cn_on_us, p_us_on_cn)
```

### Figure

```{r}
ceiling_dec <- function(x, level=1) round(x + 5*10^(-level-1), level)

military_effects$figures$mil_probs <- ggplot(data = eff_mil, aes(y = y, color = model)) + 
  geom_errorbar(aes(xmin = lb, xmax = ub), width = 0.1, linewidth = 0.5) + 
  geom_point(aes(x = md), size = 0.8) + 
  scale_x_continuous(labels = function(x){ x * 100}, 
                     limits = c(-0.05, ceiling_dec(max(eff_mil$ub)))) + 
  scale_y_continuous(name = "",
                     breaks = eff_mil$y, labels = eff_mil$y_label) +
  geom_vline(aes(xintercept = 0), linetype = "dotted", linewidth = 0.5) + 
  labs(
    title    = "Change in probability of military health presense",
    subtitle = "in response to military health presenece of other power") + 
  theme_coef(no_legend = T)

rm(ceiling_dec)
```

```{r}
png(filename = "../../Results/Figures/3-health_military-probabilities.png",
    width    = 5, 
    height   = 3,
    units    = "in",
    res      = 360)
military_effects$figures$mil_probs
dev.off()
```

### Table

```{r}
x <- eff_mil %>% 
  mutate(across(.cols = c(lb, md, ub), 
                .fns  = ~as.character(round(.x * 100, 2)))) %>% 
  mutate(out = sprintf("%s [%s, %s]", md, lb, ub)) %>%
  select(y_label, out)

military_effects$tables$mil_probs <- kbl(x = x,
                                         col.names = c("Effect", "Estimate"), 
                                         booktabs  = TRUE)
```

## Marginal effects - controls

#### GDP per capita

```{r}
m <- mods_military$us_full

x <- model.matrix(m)
x <- apply(x, 2, mean)
x[str_starts(names(x), "region")] <- 0
x["regionSub-Saharan Africa"] <- 1

b <- coef(m)

# xv <- quantile(d$un_gdppc, probs = 0.1)
xv <- quantile(d$un_gdppc, probs = 0.9)

x["un_gdppc"] <- xv
x["un_gdppc2"] <- xv^2

bx <- b["un_gdppc"] + b["un_gdppc2"] * xv

(dnorm(x %*% b) * bx * log(1.1)) * 100

```
